---
title: 'Part 1: Simulation Exercise Instructions'
author: "Yan He"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
This project is to investigate the exponential distribution in R and compare it with the Central Limit Theorem.

## Semulations
In this part we will set random seed and prepared parameters. After that we simulate exponential random variables 1000 times. Then we get 1000 results for this simulation.
```{r semulation, echo  = TRUE}
set.seed(2018)

lambda <- 0.2
n <- 40
num <- 1000

simulation_result <- matrix(rexp(n = n*num, rate = lambda), nrow = num)

```

### Mean
We calculate the mean for each row:
```{r cal_means, echo = TRUE}
simulation_means <- rowMeans(simulation_result)
```
Now we can draw a histogram plot for our semulation means and compare it to the theoretical mean of the distribution(1/lambda) which is drawn by a red line:
```{r plot1, echo = TRUE}
hist(simulation_means)
abline(v = (1/lambda), col = 'red', lwd = 5)
```

From the plot we can see that the center of simulation means is very close to the theoretical means.

### Variance
Then we calculate the variance and draw histogram to compare the simulation and true variance:
```{r cal_var, echo = TRUE}
simulation_var <- c()
for(i in 1:nrow(simulation_result)){
        s_var <- var(simulation_result[i,])
        simulation_var <- append(simulation_var, s_var)
}
hist(simulation_var)
abline(v = (1/lambda)^2, col = 'blue', lwd = 5)
```

It's obvious to see that the true variance is located in the center of our simulation vars.

### Normality
Now we try to explore whether the mean of simulation is close to normal distribution.
First, we draw a histogram plot for simulation means.Then we add a normal distribution which mean is 1/lambda and standaed deviation is 1/lambda/sqrt(n).
```{r normality, echo = TRUE}
library(ggplot2)
mean_data <- as.data.frame(simulation_means)

ggplot(data = mean_data, aes(x =simulation_means)) + 
        geom_histogram(aes(y = ..density..), 
                       bins  = 50, 
                       color = 'white',
                       alpha = 0.5, 
                       fill = 'deepskyblue4') + 
        stat_function(fun = dnorm, 
                      args = list(mean = 1/lambda,
                                  sd = (1/lambda)/sqrt(40)), 
                      color = 'lightpink',
                      size = 1)
```

It's easy to see the distribution of simulate means is very close to theoretical normal distribution from CLT.But we cannot get a precise conclusion.
We choose Kolmogorv-Smirnov test for testing the distribution of simulation means and normal distribution:

```{r ks-test, echo = TRUE}
## standard normalize
normal_sim_means <- (simulation_means - mean(simulation_means))/sd(simulation_means)

## KS test 
ks.test(normal_sim_means, 'pnorm')
```
p value is bigger than 0.05, so we cannot reject the H0 hypothesis: simulation means is normaly.
Now we can say this simulate distribution is approximately normal.
